Towards a standardized evaluation of imputation methodology

### 2022-03-30 Formatted/Formatted.tex
Write convergence issues paragraph

After deciding upon a data generating mechanism, another influential--but often overlooked--decision awaits. Simulation studies on missing data methodolology have the unique option to exclude sampling variance from their evaluations, and only use the missing data generation procedure as source of Monte Carlo variation. After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies. Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}. The only required change to simulation set-ups is to remove sampling variance from performance evaluation. A lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.

@@

Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables. It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways. Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to suboptimal advice in practice. For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that complete case analysis outperforms their imputation method(s). Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described in][\S 2.7]{fimd}, while biasing inferences in other cases. Another example of a missingness pattern that may inadverted impose assumptions on generalizations is a monotone pattern \citep[i.e. a pattern with uniformly increasing missingness proportions along variables;][]{litt20}. If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations. A final example is multivariate missingness that is generated using stepwise univariate missingness induction. The resulting incomplete data may then not have the intended or expected statistical properties \citep[e.g. skewness;][]{ampute}. This complicates any definitive conclusions about imputation method performance in relation to the data generating mechanism.

@@

\begin{table}[tb]
\begin{center}
\caption{Missingness mechanisms. [Add PMAR?]}
\label{table:mech}
\begin{tabular}{lll}
\hline
Mech. & Interpretation and consequences of the mechanism \\
\hline  
MCAR  & Missing completely at random: The probability to be missing is the same for all \\
      & cases. In other words, the missing values are missing at random and the observed \\
      & values are observed at random. \\
MAR   & Missing at random: The probability to be missing is the same within groups of \\
      & cases defined by the observed data only. The essence is that observed relations in \\
      & the data inform the missingness, such that MCAR may be assumed within the \\ 
      & observed groups. \\
MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
      & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
      & inferred from the observed data. This is considered non-ignorable missingness. \\
\hline
\end{tabular}
\end{center}
\end{table}

The missingness mechanisms that are induced in simulation studies do not necessarily match the true underlying mechanisms in incomplete empirical data, nor the mechanisms that are assumed when imputing such data. Out of the three mechanisms (see Table \ref{table:mech}), MAR missingness is usually investigated in simulations and assumed in practice. Although MCAR might not reasonably be the true missingness generating mechanism and naturally not typically assumed, the mechanism  can be very insightful in the context of imputation methodology evaluation. By contrast, MNAR

@@

[MCAR is important but often ignored:] MCAR is a necessary simulation condition for evaluating the performance of imputation procedures. Under MCAR, the statistical properties of the observed data given the missing data are known. Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not sound. Sadly, the straightforward case of MCAR is often neglected from simulation studies and focus is drawn to the evaluation of MAR mechanisms only. Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in our view may be far too simplistic.

[MAR may be spurious:] With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g. weight is made incomplete based on observed gender to emulate a situation wherin one gender is less likely to disclose their weight). These relations in the observed data may, however, be weak or non-existent. If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid. The most extreme example would be when MAR is induced from data without multivariate relations. The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random. In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p. 48]{fimd}: the missingness does not depend on the incomplete variable. This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless. %The problem of spurious MAR is generally overlooked in simulation studies

@@

[MNAR is often ignored:] Although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios. Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored during evaluation of imputation methodologies, except for those methods that are specifically targeted at non-ignorable applications.  % The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).

[Missingness type is often ignored:] Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

@@

The evaulation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e. the complete data model used to estimate a target or estimand). When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}.

[All imputation models should be congenial, but this is difficult to test and often ignored. Checks of imputation model fit are often omitted. Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others. Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies. Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]

Moreover, the performance of imputation procedures on distributional properties is often ignored in simulation studies, and even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers. For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large. While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.

% [Add segway between evaluation of imputation and actual performance measures.]
The evaluation criteria used to assess imputation performance vary from one simulator to another. This is not surprising as people from different fields could have a different focus on the problem at hand. Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification. For many outcome-driven imputation taks a single valid imputation may suffice. Measures of predictive accuracy or the RMSE (root mean squared error) of predicted values could then serve as sensible evaluation criteria.

% [Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]
% The choice of performance measures may inadvertently distort statistical properties of the imputed data. Developers often only inquire about the `accuracy' (i.e. how well can the method reproduce the original data). [Add RMSE at cell level here?]

If the goal is inference--not prediction--then the uncertainty about estimates needs to be quantified. A single imputation would then not suffice. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied. After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p. 591]{neym34}. Merely focussing on distance measures would open the door for invalid inferences.

@@

% RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods. Refer to missing data chapter. In this context, there is no talk of retrieving the `true' but unobserved values at all. The DGM is by definition design-based and there is no sampling variance. The missingness generation step is skipped and the only performance measure that can be used between methods is RMSE and maybe comparison with repect to CCA in terms of bias.]

Some general choices in the simulation set-up to consider at this stage are the simulation design and the number of simulation repetitions. The simulation design should ideally be fully factorial, i.e. varying each simulation condition against all other conditions. There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions). If there are several imputation method under evaluation, the simulation design should accomodate that each method is applied to each incomplete data set. This is in contrast to a design with as many incomplete data sets as imputation methods in every simulation repetition. Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation which makes for fairer comparisons. The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g. as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.

@@

[Use CCA as benchmark method: the imputation method should outperform CCA. Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matrix, regularization if applicable, etc.). Estimate estimand(s) using the analysis model, use correct pooling rules.]

It is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions. We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance.

@@

% = evaluation of imputation generating process
Convergence of the algorithm: Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}. The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered. 
Distributional characteristics: In practice, the distribution of the incomplete data may differ greatly from the observed data. Under anything but the MCAR assumption, this can be expected. When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for a detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}. When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.

The fit of the imputation model may be verified using posterior predictive checking. [PPC, ref Mingyang, \citep{nguy17, zhao22}.]

Plausibility of the imputed values: Plausible imputations--imputations that could be real values if they had been observed--are not a necessary condition for obtaining valid inference. However, in practice, especially when the imputer and the analyst are different persons, plausible imputations may be a desired property. One would prefer an imputation technique to yield both valid inference and plausible imputations. It should be studied if an imputation method is prone to deliver such impractical results, and if so, under what conditions. When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.

Some suggested minimal requirements for imputation methods: the absence of algorithmic non-convergence, the imputation model fits the observed parts of the data (PPC), under a MCAR mechanism the imputed values are distributed equivalent to the observed values.

@@

% sim perf and model perf
The evaluation criteria may depend on the estimand. When descriptive statistics are the goal and when statistical inference would not be of interest, the bias of the estimates would still apply, but standard errors are generally ignored. Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.

In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
 (i) Bias: Results should preferable be unbiased. However, the way bias is considered can greatly influence the interpretation of the results. For example, a negligible absolute bias for a parameter for which the true value is zero would yield infinite bias when relative bias is considered. The way bias is computed should therefore be carefully chosen and described. 
(ii) Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient. Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained. 
(iii) Interval width: The way the confidence interval is calculated should be described. Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference. However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore.

[Add RMSE at cell level and at prediction level (if applicable).]

Some suggested minimal requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis (i.e. performance after list-wise deletion).

@@

[Be explicit in choices. Describe the simulation textually, use pseudo-code, a flowchart, visualize missingness and simulation results. For example, the used missingness mechanism should be detailed, either graphically or written as a function of the data. And the procedure to generate missing data is often insufficiently described in resulting publications. ]
Fill in the following checklist (in appendix).

@@

- data types + structures (e.g. clustering/timeseries)

@@

- missingness patterns (rows = univariate/multivariate; colums = proportion)

>>>>

After deciding upon a data generating mechanism, another influential--but often overlooked--decision awaits. Simulation studies on missing data methodology have the unique option to exclude sampling variance from their evaluations, and only use the missing data generation procedure as source of Monte Carlo variation. After all, we are interested in the missingness and are not considering the noise induced by the sampling mechanism for evaluation in such studies. Therefore, it is not necessary to take sampling variation into account in order to obtain information about a method's ability to handle the missing data problem, or to objectively compare methods on their ability to correct for missingness \citep[see for a detailed discussion][]{vink14}. The only required change to simulation set-ups is to remove sampling variance from performance evaluation. A lot of published work does not consider this option, while it could have sharpened inconclusive or obfuscated results.

@@

Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables. It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways. Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to sub-optimal advice in practice. For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that list-wise deletion outperforms their imputation method(s). Such a conclusion would, unfortunately, only translate to data that adhere to the same special case \citep[described in][\S 2.7]{fimd}, while biasing inferences in other cases. Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e. a pattern with uniformly increasing missingness proportions along variables;][]{litt20}. If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations. A final example is multivariate missingness that is generated using step-wise univariate missingness induction. The resulting incomplete data may then not have the intended or expected statistical properties \citep[e.g. skewness;][]{ampute}. This complicates any definitive conclusions about imputation method performance in relation to the data generating mechanism.

@@

\begin{table}[tb]
\begin{center}
\caption{Missingness mechanisms. [Add PMAR?]}
\label{table:mech}
\begin{tabular}{lll}
\hline
Mech. & Interpretation and consequences of the mechanism \\
\hline  
MCAR  & Missing completely at random: The probability to be missing is the same for all \\
      & cases. In other words, the missing values are missing at random and the observed \\
      & values are observed at random. \\
MAR   & Missing at random: The probability to be missing is the same within groups of \\
      & cases defined by the observed data only. The essence is that observed relations in \\
      & the data inform the missingness, such that MCAR may be assumed within the \\ 
      & observed groups. \\
MNAR  & Missing not at random: The probability to be missing depends on unobserved \\
      & aspects of the data. The cause of the missingness is unknown and cannot be \\ 
      & inferred from the observed data. This is considered non-ignorable missingness \\
      & \citep[see e.g.][]{rubi76}. \\
\hline
\end{tabular}
\end{center}
\end{table}

Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions). However, the missingness mechanisms that are induced in simulation studies do not necessarily match any true missingness generating mechanisms in incomplete empirical data, nor the mechanisms that are typically assumed when imputing such data. On the one hand, there is MCAR missingness, which may be unlikely as the true missingness generating mechanism, but would yield insightful evaluations in a simulation setting. On the other hand, there is MNAR, which is arguably a quite reasonable source of many real-world missingness scenarios, but is generally ignored except for those evaluations that are specifically targeted at non-ignorable applications. A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.

% Out of the three mechanisms , random missingness (MAR) is usually generated for the evaluation of imputation methodology. Missingness mechanisms that are not random (MNAR) and mechanisms that are considered non-ignorable \citep[see e.g.][]{rubi76}, are generally ignored, except for those methods that are specifically targeted at non-ignorable applications.

% MAR missingness is usually investigated in simulations and assumed in practice. Although MCAR might not reasonably be the true missingness generating mechanism and naturally not typically assumed, the mechanism  can be very insightful in the context of imputation methodology evaluation. By contrast, MNAR

@@

MCAR is a necessary simulation condition for evaluating the performance of imputation procedures, yet often omitted. Under MCAR, the statistical properties of the observed data given the missing data are known. Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not sound. Sadly, the straightforward case of MCAR is often neglected from simulation studies and focus is drawn to the evaluation of MAR mechanisms only. Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in our view may be far too simplistic.

Although MAR missingness is often considered as simulation condition, the problem of spurious MAR is generally overlooked. With MAR missingness mechanisms, observed relations in the data are used to induce missingness during simulation (e.g. weight is made incomplete based on observed gender to emulate a situation wherein one gender is less likely to disclose their weight). These relations in the observed data may, however, be weak or non-existent. If MAR is induced based on weaker relations in the data, claims for a method's applicability to situations where the missingness is random become less valid. The most extreme example would be when MAR is induced from data without multivariate relations. The inferential implications of the missingness would then mimic those of MCAR, even though the missingness is random. In fact, this would amount to one of the special cases under which complete case analysis would be more efficient than multiple imputation \citep[see e.g.][p. 48]{fimd}: the missingness does not depend on the incomplete variable. This property might be useful in practice, but considering it as a condition to evaluate performance under MAR missingness is useless.

Inducing a MAR mechanism comes with another issue, namely the type or functional form of the missingness model. Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

@@

Finally, although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios. As stated, this mechanism is not usually included in evaluations.  % The model--if any--that is used to generate the missingness is usually assumed to be random (MAR) or completely random (MCAR).

@@

The evaluation criteria used to assess imputation performance vary from one simulator to another. This is not surprising as people from different fields could have a different focus on the problem at hand. There are, however, some over-arching issues with assessing imputation method performance. In the first place, evaluations of the imputation-generating process (e.g., an iterative imputation algorithm) are generally left out of simulation study results. In the second place, there are performance measures pitfalls to consider.

Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations. As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated. Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{fimd} is neither efficient nor failproof. As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method. Moreover, failing to produce stable imputations is not an exclusive quality iterative aimputation algorithms. For example if a method does not yield any imputations at all in a certain simulation repetition (e.g. due to over-parameterization), this can be seen as failure in the imputation-generating process too.

The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models (i.e. the complete data model used to estimate a target or estimand). When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}. Congeniality is difficult to diagnose, but ignoring it may yield sub-optimal imputations. Methods to asses the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}, which is an infeasible endeavor in many simulation studies. % [All imputation models should be congenial, but this is difficult to test and often ignored. Checks of imputation model fit are often omitted. Ignoring congeniality may yield sub-optimal imputations, which may disadvantage certain methods compared to others. Same goes for convergence: FCS is iterative and requires algorithmic convergence, but this is typically evaluated through visual inspection which is unfeasible in simulation studies. Problems with convergence may be even more severe: some methods do not produce imputations at all for certain conditions, while at other times the imputation algorithm is cut off too soon, and performance will be under-estimated.]  
Therefore, the performance of imputation procedures on distributional properties is often ignored in simulation studies. Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers. For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large. While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny.

% [Add segway between evaluation of imputation and actual performance measures.]
% [Add that simulation aim/target determines which performance measures are relevant to consider, and that the choice of analysis model/estimand can influence the values of these metrics (e.g., focusing on univariate estimates only may favor ad hoc methods).]
% The choice of performance measures may inadvertently distort statistical properties of the imputed data. Developers often only inquire about the `accuracy' (i.e. how well can the method reproduce the original data). [Add RMSE at cell level here?]

If the goal is inference--not prediction--then the uncertainty about estimates needs to be quantified. A single imputation would then not suffice. The aim of multiple imputation is not to reproduce the data, but to allow for obtaining valid inference given that the data are incomplete. This means that, given the framework provided by \citet{rubi87}, statistical properties such as bias, confidence intervals, and the coverage rate of the confidence intervals should be studied. After all, the 95\% confidence interval should contain the `true' value at least 95 out of 100 times \citep[][p. 591]{neym34}. Merely focusing on distance measures would open the door for invalid inferences.

Some simulators may focus on retrieving the unobserved values, while others study the performance in the context of obtaining efficient predictions or classification. For many outcome-driven imputation task a single valid imputation may suffice. Measures of predictive accuracy or the RMSE (root mean squared error) of predicted values could then serve as sensible evaluation criteria.

@@

% RMSE of the predicted values after imputation and estimation to say something about the comparative performance of the methods. Refer to missing data chapter. In this context, there is no talk of retrieving the `true' but unobserved values at all. The DGM is by definition design-based and there is no sampling variance. The missingness generation step is skipped and the only performance measure that can be used between methods is RMSE and maybe comparison with respect to CCA in terms of bias.]

Some general choices in the simulation set-up to consider at this stage are the simulation design and the number of simulation repetitions. The simulation design should ideally be fully factorial, i.e. varying each simulation condition against all other conditions. There are, for example, known interaction effects between missingness mechanisms and missingness patterns (the validity of the assumed missingness mechanism becomes increasingly important with higher missingness proportions). If there are several imputation method under evaluation, the simulation design should accommodate that each method is applied to each incomplete data set. This is in contrast to a design with as many incomplete data sets as imputation methods in every simulation repetition. Applying all methods to the same incomplete data set is computationally convenient, and minimizes unnecessary variation which makes for fairer comparisons. The number of simulation repetitions may be informed by the required level of precision in the simulation study \citep[e.g. as determined from a maximum tolerable level of uncertainty in terms of a performance measure's Monte Carlo error][]{morr18}.

@@

The imputation method(s) under evaluation should be applied to the generated incomplete data. It is important to choose the imputation model parameters carefully (e.g., number of imputations, number of iterations, regularization, etc.). After imputation, an analysis model may be fitted to estimate the estimand(s).

%[Use CCA as benchmark method: the imputation method should outperform CCA. Choose imputation model parameters carefully (number of imputations, number of iterations if applicable, predictor matrix, regularization if applicable, etc.). Estimate estimand(s) using the analysis model, use correct pooling rules.]

Apart from the imputation method(s), list-wise deletion should be considered as missing data method too. We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating multiple imputation performance. List-wise deletion may therefore serve as a benchmark method: out-performing it should be a minimal requirement for the imputation method(s). %it is wise to evaluate the performance of complete case analysis (aka list-wise deletion) in all simulated conditions.

@@

[Some suggested minimal requirements for imputation methods: the absence of algorithmic non-convergence, the imputation model fits the observed parts of the data (PPC), under a MCAR mechanism the imputed values are distributed equivalent to the observed values.]

% = evaluation of imputation generating process
The absence of non-convergence in the imputation models is a minimal requirement for imputation methods. Most contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, where some algorithms are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}. The convergence of all iterative algorithms should always be considered and if non-convergence is suspected, the inference resulting from the imputations should not be considered.

[Distributional characteristics] In practice, the distribution of the incomplete data may differ greatly from the observed data. Under anything but the MCAR assumption, this can be expected. When evaluating imputations, the distributional shapes should be checked and diagnostic evaluations should be performed \citep[see][for a detailed overview of diagnostic evaluation for multivariate imputations]{abayomi2008diagnostics}. When anomalies are found, and if the imputation method is valid, there should be an explanation, especially in the controlled environment of a properly executed simulation study.

The fit of the imputation model may be verified with the help of a posterior predictive check \citep[PPC][]{nguy17, zhao22}. A PPC in the context imputation methodology could be the over-imputation of observed data values. The reasoning is that if the over-imputed values have  equivalent statistical properties to the true data values, the imputation model fits the observed part of the incomplete data well. We might then assume that the imputation model would be able to produce good imputations for missing values as well.

[Plausibility of the imputed values] Plausible imputations--imputations that could be real values if they had been observed--are not a necessary condition for obtaining valid inference. However, in practice, especially when the imputer and the analyst are different persons, plausible imputations may be a desired property. One would prefer an imputation technique to yield both valid inference and plausible imputations. It should be studied if an imputation method is prone to deliver such impractical results, and if so, under what conditions. When evaluating imputation routines, the evaluator should mention whether the routine is prone to deliver implausible values.

@@

[Some suggested minimal requirements for imputation methods: the method yields valid inference under MCAR, the method performs equally well or better than complete case analysis (i.e. performance after list-wise deletion).]

% sim perf and model perf
[The evaluation criteria may depend on the estimand. When descriptive statistics are the goal and when statistical inference would not be of interest, the bias of the estimates would still apply, but standard errors are generally ignored. Chances are that for those who focus on descriptive statistical applications, multiple imputation would not be the mode of choice.]

In general, we would say that each multiple imputation routine should be evaluated on at least the following points:
 (i) Bias: Results should preferable be unbiased. However, the way bias is considered can greatly influence the interpretation of the results. For example, a negligible absolute bias for a parameter for which the true value is zero would yield infinite bias when relative bias is considered. The way bias is computed should therefore be carefully chosen and described. 
(ii) Coverage: Coverage of a 95\% interval should in theory be $\geq 95$, where a coverage rate of 95\% would be most efficient. Under-coverage (when estimation is too liberal) may be an indication of invalid inference, while over-coverage (when estimation is too conservative) tells us that efficiency could still be gained. 
(iii) Interval width: The way the confidence interval is calculated should be described. Wider intervals are associated with more uncertainty and the more narrow interval that is still properly covered indicates a sharper inference. However, inference from a wider interval that is properly covered is to be considered more valid when compared to a more narrow interval that is not properly covered anymore. 
% [Add RMSE at cell level and at prediction level (if applicable).]

@@

[Be explicit in choices. Describe the simulation textually, use pseudo-code, a flowchart, visualize missingness and simulation results. For example, the used missingness mechanism should be detailed, either graphically or written as a function of the data. And the procedure to generate missing data is often insufficiently described in resulting publications. Fill in the following checklist (in appendix).]

@@

- data types + structures (e.g. clustering/time-series)

@@

- missingness patterns (rows = univariate/multivariate; columns = proportion)
